Stellar Radiation is Critical for Regulating Star Formation and Driving Outflows in Low Mass Dwarf Galaxies

### 2018-09-19 ms.tex
final edits for acceptance

\begin{abstract}
%
%
% NOTE: Overleaf includes abstract words in word count in text... subtract
%       this from the total.
%
%
Effective stellar feedback is used in models of galaxy formation to 
%regulate star formation and 
drive realistic galaxy evolution. Models typically include energy injection from supernovae as the dominant form of stellar feedback, often in some form of sub-grid recipe. However, it has been recently suggested that pre-SN feedback (stellar winds or radiation) is necessary in high-resolution simulations of galaxy evolution to properly regulate star formation and properties of the interstellar medium (ISM). However, following these processes is computationally challenging, 
%Following stellar winds and ionizing radiation is computationally challenging, 
so many prescriptions model this feedback approximately, accounting for the local destruction of dense gas clouds around newly formed stars in lieu of a full radiative transfer calculation. In this work we examine high resolution simulations (1.8~pc) of an isolated dwarf galaxy with detailed stellar feedback tracked on a star-by-star basis. By following stellar ionizing radiation with an adaptive ray-tracing radiative transfer method, we test its importance in regulating star formation and driving outflows in this galaxy. We find that including ionizing radiation reduces the star formation rate (SFR) by over a factor of 5, and is necessary in creating the ISM conditions needed for supernovae to drive significant outflows. We find that a localized approximation for radiation feedback is sufficient to regulate the SFR on short timescales, but is unable to drive significant outflows. Short and long range radiation effects are both important in driving the evolution of our \textbf{low-metallicity, low-mass dwarf galaxy. Generalization of these results to more massive galaxies would be a valuable avenue of future research.}
\end{abstract}

@@

We additionally present a simulation run without any stellar ionizing radiation (``noRT'') and a second simulation that includes ionizing radiation, but deletes all photons that travel more than 20~pc from their source (``shortrad''). This second simulation tests the relative importance of short-range vs. long-range effects of stellar ionizing radiation as a form of feedback. Although this is still more accurate than an approximate method, this is meant to function similarly to methods that include only the local effects of stellar radiation feedback. Each of our three simulations is identical up until the formation of the first star particles.\footnote{We note that the stars from the very first star formation event in each run (21 star particles with a total mass of 114 M$_{\odot}$) are the same. Of these, one particle emits ionizing radiation ($M_* > 8 $~M$_{\odot}$). \textbf{The sampled stellar masses differ across runs after this point. Stochastic effects from differences in IMF sampling may be important in the regime of low mass, low SFR dwarf galaxies. Multiple re-simulations of identical initial conditions may result in significant scatter in their final properties \citep{Keller2018}. We find our main conclusions are likely to be insensitive to these stochastic variations, as we find that the total radiation output across multiple re-samplings of a fixed star formation history is small after the IMF is fully sampled ($<50$~Myr).}}

\textbf{As shown in Paper I, the maximum densities reached in these simulations is below 10$^{3}$~cm$^{-3}$. Our inability to resolve the high densities in star forming regions ($\sim 10^5$~cm$^{-3}$) means that we likely underestimate initial photon absorption and overestimate the initial long-range effects of newly formed stars. However, in the Milky Way, newly formed O stars have been observed to spend no more than  10 -- 20\% of their main sequence lifetimes embedded in ultracompact HII regions within dense molecular clouds \citep{WoodChurchwell1989}. This short dispersal timescale is confirmed by high resolution simulations of massive star formation \citep[e.g.][]{Peters2010,Dale2014,Kim2018}. Therefore, we do not consider our neglect of this phase to be likely to lead to substantial errors at the galactic scale.}

@@

\begin{figure*}
\centering
\includegraphics[width=0.99\linewidth]{DD0136_fiducial_shortrad_nort}
\caption{Edge-on slices of each simulation showing gas number density (left), temperature (middle), and hydrogen ionization fraction (right) 17~Myr after the formation of the first star in each run. Each panel is 4 kpc x 4 kpc. \textbf{The full movie of this evolution is included for further clarification of the comparison between this figure and Fig.~\ref{fig:panel2}}}
\label{fig:panel1}
\end{figure*}

@@

\textbf{These results show that long-range ionization effects are an important consideration in models of stellar feedback. However, it warrants further study as to how much this effect can be accounted for without resorting to full radiative transfer calculations. Approximate, Str{\"o}mgren-like feedback models which allow for ionization far from a single source \citep[e.g.][]{Hopkins2018} may be sufficient to capture the effects shown here. Fully localized methods, or methods which set a maximum for the ionization radius may underestimate galactic wind properties in dwarf galaxies. In addition, both methods are mass biased \citep[see the discussion in ][]{Hu2017}, preferentially over-ionizing dense gas that would otherwise be missed by photons leaking through channels carved through lower-density gas. These remaining uncertainties motivate a continued examination of the feedback prescriptions adopted in high resolution simulations of galaxy evolution.}

@@

\software{astropy \citep{astropy} matplotlib \citep{matplotlib}, Numpy \citep{numpy}, SciPy \citep{scipy}, IPython \citep{Ipython}, yt \citep{yt}, Enzo \citep{Enzo2014}, Grackle \citep{GrackleMethod}, CLOUDY \citep{cloudy}, HDF5}

>>>>

\begin{abstract}
%
%
% NOTE: Overleaf includes abstract words in word count in text... subtract
%       this from the total.
%
%
Effective stellar feedback is used in models of galaxy formation to drive realistic galaxy evolution. Models typically include energy injection from supernovae as the dominant form of stellar feedback, often in some form of sub-grid recipe. However, it has been recently suggested that pre-SN feedback (stellar winds or radiation) is necessary in high-resolution simulations of galaxy evolution to properly regulate star formation and properties of the interstellar medium (ISM). Following these processes is computationally challenging, so many prescriptions model this feedback approximately, accounting for the local destruction of dense gas clouds around newly formed stars in lieu of a full radiative transfer calculation. In this work we examine high resolution simulations (1.8~pc) of an isolated dwarf galaxy with detailed stellar feedback tracked on a star-by-star basis. By following stellar ionizing radiation with an adaptive ray-tracing radiative transfer method, we test its importance in regulating star formation and driving outflows in this galaxy. We find that including ionizing radiation reduces the star formation rate (SFR) by over a factor of 5, and is necessary to produce the ISM conditions needed for supernovae to drive significant outflows. We find that a localized approximation for radiation feedback is sufficient to regulate the SFR on short timescales, but does not allow significant outflows. Short and long range radiation effects are both important in driving the evolution of our low-metallicity, low-mass dwarf galaxy. Generalizing these results to more massive galaxies would be a valuable avenue of future research.
\end{abstract}

@@

We additionally present a simulation run without any stellar ionizing radiation (``noRT'') and a second simulation that includes ionizing radiation, but deletes all photons that travel more than 20~pc from their source (``shortrad''). This second simulation tests the relative importance of short-range vs. long-range effects of stellar ionizing radiation as a form of feedback. Although this is still more accurate than an approximate method, this is meant to function similarly to methods that include only the local effects of stellar radiation feedback. Each of our three simulations is identical up until the formation of the first star particles.\footnote{We note that the stars from the very first star formation event in each run (21 star particles with a total mass of 114 M$_{\odot}$) are the same. Of these, one particle emits ionizing radiation ($M_* > 8 $~M$_{\odot}$). The sampled stellar masses differ across runs after this point. Stochastic effects from differences in IMF sampling may be important in the regime of low mass, low SFR dwarf galaxies. Multiple re-simulations of identical initial conditions may result in significant scatter in their final properties \citep{Keller2018}. We find our main conclusions are likely to be insensitive to these stochastic variations, as we find that the total radiation output across multiple re-samplings of a fixed star formation history is small after the IMF is fully sampled.}

As shown in Paper I, the maximum densities reached in these simulations is below 10$^{3}$~cm$^{-3}$. Our inability to resolve the high densities in star forming regions ($\sim 10^5$~cm$^{-3}$) means that we likely underestimate initial photon absorption and overestimate the initial long-range effects of newly formed stars. However, in the Milky Way, newly formed O stars have been observed to spend no more than  10 -- 20\% of their main sequence lifetimes embedded in ultracompact HII regions within dense molecular clouds \citep{WoodChurchwell1989}. This short dispersal timescale agrees with high resolution simulations of massive star formation \citep[e.g.][]{Peters2010,Dale2014,Kim2018}. Therefore, we do not consider our neglect of this phase to be likely to lead to substantial errors at the galactic scale.

@@

\begin{figure*}
\centering
\includegraphics[width=0.99\linewidth]{DD0136_fiducial_shortrad_nort}
\caption{Edge-on slices of each simulation showing gas number density (left), temperature (middle), and hydrogen ionization fraction (right) 17~Myr after the formation of the first star in each run. Each panel is 4 kpc x 4 kpc. The full movie of this evolution is included for further clarification of the comparison between this figure and Fig.~\ref{fig:panel2}.}
\label{fig:panel1}
\end{figure*}

@@

These results show that long-range ionization effects are an important consideration in models of stellar feedback. However,
further study is warranted of how this effect can be approximated without resorting to full radiative transfer calculations. Approximate, Str{\"o}mgren-like feedback models that allow for ionization far from a single source \citep[e.g.][]{Hopkins2018} may be sufficient to capture the effects shown here. Fully localized methods, or methods which set a maximum for the ionization radius may underestimate galactic wind properties in dwarf galaxies. In addition, both methods are mass biased \citep[see the discussion in ][]{Hu2017}, preferentially over-ionizing dense gas that would otherwise be missed by photons leaking through channels carved through lower-density gas. These remaining uncertainties motivate a continued examination of the feedback prescriptions adopted in high resolution simulations of galaxy evolution.

@@

\software{astropy \citep{astropy} matplotlib \citep{matplotlib}, Numpy \citep{numpy}, SciPy \citep{scipy}, IPython \citep{Ipython}, yt \citep{yt}, Enzo \citep{Enzo2014}, Grackle \citep{GrackleMethod}, CLOUDY \citep{cloudy}, HDF5 \citep{HDF5}}
